Temperature and moisture effects on greenhouse gas emissions from deep active-layer boreal soils

### 2016-11-19 manuscript/response_to_reviewer_comments.tex

\medskip
{\it 1.6. Studies have shown that moisture can have a weaker effect on temperature sensitivity early on during an incubation experiment, in the presence of more labile C. This relative to the effect on moisture on the Q-10 of cumulative respiration, reflecting slow turning over C - this could be an interesting analysis to include here, and would help to assess how }

This is an interesting suggestion, and an analysis we are happy to include. We do not observe any evidence for changes in CO2 moisture sensitivity with time, and weak changes in CO2 temperature sensitivity; CH4 emissions show a weak decline in moisture sensitivity with time. This is now discussed in the text. {\bf See new lines LINES.}

\medskip
{\it 1.7. How do your results in terms of temperature and moisture sensitivity (especially under drought conditions) scale with Alaskan climate change predictions from modelers? How does it compare with deep soils incubations (mineral soils) from the Arctic, and from boreal ecosystems? }

The first question is similar to a point raised by Referee \#2 (comment 2.3). We have added this as a paragraph in the discussion, noting e.g. observable anthropogenic influences on high-latitude precipitation, drier and warmer conditions in boreal Eurasia, and growing season length increases in interior Alaska with no increase in precipitation. {\bf See new lines LINES.}

@@

{\it 2.2. It seems like this paper was published as a discussion paper before Schädel et al. 2016 was published and hence a discussion of the meta-analysis was not possible but should be addressed in the revisions. }

This point was also made by Referee \#1 (comment 1.18). Yes, the fact that we didn't cite the Schädel et al. (2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted. We have significantly expanded the discussion on this point, comparing our results to Schädel et al. (2016) in depth, particularly their findings of higher aerobic than anaerobic respiration; respiration dominance of CO2 versus CH4; and Q10 values. {\bf See new lines LINES.}

\medskip
{\it 2.3. The importance of the results would be more obvious if the discussion also contained an upscaling or circumpolar aspect of drought in the Arctic. It would be useful to have some discussion about the area that is expected to be most affected by drought. This is important as changes in temperature will affect most of the Arctic, whereas drought effects or dry soils will occur more locally. }

We have added this as a paragraph in the discussion, noting e.g. observable anthropogenic influences on high-latitude precipitation, drier and warmer conditions in boreal Eurasia, and growing season length increases in interior Alaska with no increase in precipitation. {\bf See new lines LINES.}

### 2016-11-19 manuscript/cpcrw_manuscript.Rmd
Upscaling paragraph (reviewer comments 1.7 and 2.3)

---
output: word_document
---
[//]: This Rmarkdown file generates the Biogeosciences manuscript

@@

```{r setup, echo=FALSE, fig.keep='none', message=FALSE, results='hide'}
library(dplyr, warn.conflicts = FALSE)

# Run the main analysis script, which will leave figures, tables, and data 
# in memory for us. Need to temporarily change the working directory to do so.
olddir <- getwd()
setwd("../")
source("R/5-analyze.R")
setwd(olddir)
```

@@

A number of synthesis studies have produced robust results documenting the dynamics and C feedback potential of Arctic and boreal soils; comparing to these results is useful because although the response of soil biota to stresses such as drought tends to differ between soil types, organisms, and vegetation, it is often broadly similar across biomes and climatic conditions {Manzoni, 2012 #3545}. Using two meta-analyses of aerobic and anaerobic permafrost soil incubations, Schädel et al. {, 2016 #4444} showed that C release was highly sensitive to temperature, and that soils released far more (220-520%) C under aerobic conditions. Our incubation was fully aerobic, but its results are consistent with the conclusion that respiration in the form of CO~2~ is likely to dominate the high latitude C feedback, and that aerobic soils, and the conditions under which currently waterlogged soils may drain, deserve particular attention. In terms of absolute flux rates, in a pan-Arctic synthesis of anaerobic soil incubations, Treat et al. {, 2015 #4217} reported mean CO~2~ rates of 47 (all mineral soils) and 101 (for 20-100 cm soils) µg C-CO~2~ g C^-1^ day^-1^, somewhat lower than our aerobic incubation results. Treat et al. {, 2014 #4213} also found CO~2~ and CH~4~ emissions to be strongly correlated with temperature and moisture based on an incubation of Alaskan peats. Whether climate change makes northern regions wetter or driers is thus a critical factor affecting the quantity and form of C release.

>>>>

---
output:
  html_document: default
  word_document: default
---
[//]: This Rmarkdown file generates the Biogeosciences manuscript
[//]: Built using rmarkdown 1.1, RStudio 1.0.44, and R 3.3.1
[//]: November 23, 2016

@@

```{r setup, echo=FALSE, fig.keep='none', message=FALSE, results='hide'}
# Run the main analysis script, which will leave figures, tables, 
# and data in memory for us to access. 
# Need to temporarily change the working directory to do so.
olddir <- getwd()
setwd("../")
source("R/5-analyze.R")
setwd(olddir)
```

@@

A number of synthesis studies have produced robust results documenting the dynamics and C feedback potential of Arctic and boreal soils; comparing to these results is useful because although the response of soil biota to stresses such as drought tends to differ between soil types, organisms, and vegetation, it is often broadly similar across biomes and climatic conditions {Manzoni, 2012 #3545}. Using two meta-analyses of aerobic and anaerobic permafrost soil incubations, Schädel et al. {, 2016 #4444} showed that C release was highly sensitive to temperature, and that soils released far more (220-520%) C under aerobic conditions. Our incubation was fully aerobic, but its results are consistent with the conclusion that respiration in the form of CO~2~ is likely to dominate the high latitude C feedback, and that aerobic soils, and the conditions under which currently waterlogged soils may drain, deserve particular attention. In terms of absolute flux rates, Treat et al. {, 2015 #4217} reported mean CO~2~ rates of 47 (all mineral soils) and 101 (for 20-100 cm soils) µg C-CO~2~ g C^-1^ day^-1^ from a pan-Arctic synthesis of anaerobic soil incubations, somewhat lower than our aerobic incubation results. Treat et al. {, 2014 #4213} also found CO~2~ and CH~4~ emissions to be strongly correlated with temperature and moisture based on an incubation of Alaskan peats. Whether climate change makes northern regions wetter or driers is thus a critical factor affecting the quantity and form of C release.

The drought treatment imposed in this experiment reduced soil C fluxes by `r pn(drought_effect["20", "Reduction"], 0)`% - `r pn(drought_effect["4", "Reduction"], 0)`%. The importance of this result depends, in part, on the spatial extent and intensity of precipitation changes across the boreal and Arctic this century. There is a detectable anthropogenic influence in high latitude precipitation {Wan, 2015 #4478}, and its changes in which exhibit high spatial variability: drier and warmer conditions in boreal Eurasia {Buermann, 2014 #3992}, for example, but growing season length increases in interior Alaska with no increase in precipitation {Wendler, 2009 #4477}. This spatial variability will interact with permafrost thaw dynamics to produce a complex patchwork of soil moisture changes {Zhang, 2012 #4479}{Watts, 2012 #4505}. This high uncertainty in this area makes it all the more important to understand the interactive effects of soil moisture and temperature on decomposition and GHG emissions {Sierra, 2015 #4203}.
